\documentclass[a4paper]{scrreprt}

% Uncomment to optimize for double-sided printing.
% \KOMAoptions{twoside}

% Set binding correction manually, if known.
% \KOMAoptions{BCOR=2cm}

% Localization options
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

% Quotations
\usepackage{dirtytalk}

% Enhanced verbatim sections. We're mainly interested in
% \verbatiminput though.
\usepackage{verbatim}

% Automatically remove leading whitespace in lstlisting
\usepackage{lstautogobble}

% PDF-compatible landscape mode.
% Makes PDF viewers show the page rotated by 90°.
\usepackage{pdflscape}

% Advanced tables
\usepackage{array}
\usepackage{tabularx}
\usepackage{longtable}

% Fancy tablerules
\usepackage{booktabs}

% Graphics
\usepackage{graphicx}

% Current time
\usepackage[useregional=numeric]{datetime2}

% Float barriers.
% Automatically add a FloatBarrier to each \section
\usepackage[section]{placeins}

% Custom header and footer
\usepackage{fancyhdr}

\usepackage{geometry}
\usepackage{layout}

% Math tools
\usepackage{mathtools}
% Math symbols
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{amsthm}

\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}

% General symbols
\usepackage{stmaryrd}

\DeclarePairedDelimiter\abs{\lvert}{\rvert}

% Indistinguishable operator (three stacked tildes)
\newcommand*{\diffeo}{% 
  \mathrel{\vcenter{\offinterlineskip
  \hbox{$\sim$}\vskip-.35ex\hbox{$\sim$}\vskip-.35ex\hbox{$\sim$}}}}

% Bullet point
\newcommand{\tabitem}{~~\llap{\textbullet}~~}

\pagestyle{plain}
% \fancyhf{}
% \lhead{}
% \lfoot{}
% \rfoot{}
% 
% Source code & highlighting
\usepackage{listings}

% Coloured boxes!
\usepackage[most]{tcolorbox}
\newtcolorbox{library}[2][]{
	enhanced,
	sharp corners,
	coltitle=black, % title colour
	colbacktitle=black!10!white, % title bg colour
	halign title=center, % title align
	toptitle=1mm, % Top/bottom additional spacing for title
	bottomtitle=1mm,
	fonttitle=\ttfamily,
	colback=white, % body bg colour
	fontupper=\ttfamily,
	title=#2,#1
}

\newtcolorbox{boxcomment}[2][]{
	enhanced,
	colframe=white, % frame colour
	colbacktitle=white, % title bg colour
	halign=center, % body align
	colback=white, % body bg colour
	fonttitle=\ttfamily,
	fontupper=\ttfamily,
	title=#2,#1
}

% SI units
\usepackage[binary-units=true]{siunitx}
\DeclareSIUnit\cycles{cycles}

% Convenience commands
\newcommand{\mailsubject}{41100 - Distributed algorithms - Series 3}
\newcommand{\maillink}[1]{\href{mailto:#1?subject=\mailsubject}
                               {#1}}

% Should use this command wherever the print date is mentioned.
\newcommand{\printdate}{\today}

\subject{41102 - Distributed algorithms}
\title{Series 3}

\author{Michael Senn \maillink{michael.senn@students.unibe.ch} - 16-126-880}

\date{\printdate}

% Needs to be the last command in the preamble, for one reason or
% another. 
\usepackage{hyperref}


\begin{document}
\maketitle


\setcounter{chapter}{2}

\chapter{Series 3}

\section{Relations among failure detectors}

Given a non-perfect failure detector $NP$ with strong accuracy and weak
completness, a perfect failure detector $P$ can be implemented as follows. This
utilizes the fact that, if a process $p$ running $NP$ detects a process $p'$ to
have crashed, due to the strong accuracy property of $NP$, $p'$ is guaranteed
to actually have crashed.

\begin{library}{Implementing $P$ using $NP$}
  \begin{lstlisting}[mathescape=true,autogobble=true,breaklines=true]
    UPON init:
      alive := $\Pi$
      detected := {}

    UPON <NP.crash | p>:
      $\text{detected} := \text{detected} \cup \{p\}$
      $\text{alive} := \text{alive} \setminus \{p\}$

      SEND [p crashed] to all $p' \in \Pi \setminus \{self\}$
      TRIGGER <P.crash | p>

    UPON receive [p crashed] from $p'$:
      $\text{detected} := \text{detected} \cup p$
      $\text{alive} := \text{alive} \cup p$
      TRIGGER <P.crash | p>
  \end{lstlisting}
\end{library}

\section{Perfect Failure Detector}

For conciseness we assume that we have access to a slightly extended timer
object, with the following properties:
\begin{itemize}
  \item Each timer carries a reference to a process $p$, which is emitted as a
    parameter in its timeout event.
  \item Each such timer can be acessed via its parameter $p$.
  \item Each timer can be reset - that is resetting its countdown to the value
    it was started with.
\end{itemize}

With this, a perfect failure detector based on unidirectional heartbeats can be
implemented as follows, with an identical interface to the perfect failure
detector presented in the lecture.

\begin{library}{Implementing $P$ using heartbeats}
  \begin{lstlisting}[mathescape=true,autogobble=true,breaklines=true]
    UPON init:
      alive := $\Pi$
      detected := {}

      for $p \in (\Pi \setminus \{self\})$:
        // Per process timer to track lifetime of remote processes
        start_timer($\Delta + \Phi + 1$, p)

      // Start timer to periodically send own heartbeats, and send initial heartbeat
      start_timer{$\Delta + \Phi$, self)
      SEND [HEARTBEAT] to all $p' \in \text{alive} \setminus \{\text{self}\}$
      

    UPON <timeout | p>:
      if p == self:
	reset_timer(self)
	SEND [HEARTBEAT] to all $p' \in \text{alive} \setminus \{\text{self}\}$
      else:
        $\text{detected} := \text{detected} \cup \{p\}$
        $\text{alive} := \text{alive} \setminus \{p\}$

        TRIGGER <P.Crash | p>

    UPON receive [HEARTBEAT] from p:
      // Restart timer tracking lifetime of $p$
      reset_timer(p)
  \end{lstlisting}
\end{library}

As every process sends a heartbeat in the beginning, as well as every $\Delta +
\Phi$, with each heartbeat taking at most $\Delta + \Phi$ to arrive and be
processed, not receiving any heartbeat from any process for $\Delta + \Phi + 1$
guarantees that it is not alive.

\section{Quorum systems}

\subsection{Singleton}

If the sole process $p \in \Pi$ is faulty, there exists no singleton quorum
system $\mathcal{Q}$, as its sole quorum would have to contain $p$. Hence the
minimum and maximum number of tolerable failed process is $0$.

\subsection{Majority}

As each quorum of the majority quorum system has a cardinality of
$\ceil*{\frac{n + 1}{2}}$, there must clearly be at most $n - \ceil*{\frac{n +
1}{2}}$ faulty processes in order to get at least one quorum satisfying the
requirement. For $n \in \mathbb{N}$ this is equal to $\floor*{\frac{n -
1}{2}}$.

The choice of these faulty processes does not matter, as such the minimum and
maximum number of tolerable faulty processes is equal.

\subsection{Grid}

The quorum of the grid system with the lowest cardinality is the one where
exclusively the processes of the last row are selected, with a cardinality of
$k$. In such a system there must be at least $k$ correct, and hence at most
$k^2 - k$ faulty processes.

Similarly the quorum with the highest cardinality is the one where the first
row of processes, and one process of each subsequent row, is selected. Its
cardinality is given as $k + k - 1$, in which case there are at most $k^2 - 2k
+ 1$ faulty tolerable processes.

As such the quorum system can tolerate at least $k^2 - 2k + 1$, and at most
$k^2 - k$ faulty processes.

\end{document}
